Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-28 paper/intro.tex
Add code contract to intro and fix the outline

Two cyber-resiliency transformations are discussed: (1) the insertion
of a filter to prevent malformed data from a malicious actor from
being propagated to downstream components, and (2) the insertion of a
monitor to detect (and alert) unexpected behaviors arising from
untrusted components. These transformations not only change the
architecture of the model by adding in new components; they also
generate a formal specification for the behavior of the inserted
high-assurance components in the AGREE language. Those specifications
are sufficient for model checking to prove that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

Another novel aspect of the approach is the synthesis of the AGREE
specifications for the high-assurance components to CakeML, a verified
compiler implementation for the functional programming language
ML \cite{cakeml}. This paper describes in detail the synthesis path
from specifications to CakeML code, providing a formal framework in
which to argue correctness. CakeML then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the CakeML source code is exactly
preserved in the final binaries). Assuming that the execution schedule
of the deployed cyber-hardened system is as intended by the AADL
model, and that the HAMR-generated communication fabric delivers
messages between components as expected, the AGREE model checking
results then hold for the deployed system, \ie, the system will detect
and prevent the indicated cyber-vulnerabilities over all possible
finite inputs. Preliminary work has shown how to lift this result to
infinite input traces as these systems are inherently reactive and
intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

@@

The approach is illustrated by a simple example in
Section~\ref{sec:example}. Contract specification in the AGREE
language is presented in Section~\ref{sec:agree}, followed by a
description of the synthesis pathway in Section~\ref{sec:synthesis}. A
case study applying these transformations to an Unmanned Aerial
Vehicle (UAV) system that uses the Air Force Research Laboratory's
OpenUxAS services for route planning is presented in
Section~\ref{sec:case-study}. Here the transforms add filters to guard
against malformed input and monitors to guard against ground station
spoofing and malicious flight plans from OpenUxAS. The case study
system is significantly more complex than the simple example and shows
the viability of the modeling approach to a full-scale industrial
design.

>>>>

Two cyber-resiliency transformations are discussed: (1) the insertion
of a filter to prevent malformed data from a malicious actor from
being propagated to downstream components, and (2) the insertion of a
monitor to detect (and alert) unexpected behaviors arising from
untrusted components. These transformations not only change the
architecture of the model by adding in new components; they also
generate a formal specification in the form of a \emph{code contract} in the AGREE language for their intended behavior. 
A code contract describes the computation of a high-assurance component and not just what it computes.
These code contracts
are sufficient for model checking to prove that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

Another novel aspect of the approach is the synthesis of the AGREE
code contracts for the high-assurance components to CakeML, a verified
compiler implementation for the functional programming language
ML \cite{cakeml}. This paper describes in detail the synthesis path
from code contracts to CakeML code, providing a formal framework in
which to argue correctness. CakeML then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the CakeML source code is exactly
preserved in the final binaries). Assuming that the execution schedule
of the deployed cyber-hardened system is as intended by the AADL
model, and that the HAMR-generated communication fabric delivers
messages between components as expected, the AGREE model checking
results then hold for the deployed system, \ie, the system will detect
and prevent the indicated cyber-vulnerabilities over all possible
finite inputs. Preliminary work has shown how to lift this result to
infinite input traces as these systems are inherently reactive and
intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

@@

The approach is illustrated by a simple example in
\secref{sec:example}. AGREE verification and specification in the AGREE
language are presented in \secref{sec:agree} and \secref{agree-semantics}.
Code contracts are defined in \secref{sec:code-contracts} with a method
for testing code contracts in the AGREE framework before synthesis given
\secref{sec:testing}.
The synthesis pathway is covered in Section~\ref{sec:synthesis}. A
case study applying these transformations to an Unmanned Aerial
Vehicle (UAV) system that uses the Air Force Research Laboratory's
OpenUxAS services for route planning is presented in
Section~\ref{sec:case-study}. Here the transforms add filters to guard
against malformed input and monitors to guard against ground station
spoofing and malicious flight plans from OpenUxAS. The case study
system is significantly more complex than the simple example and shows
the viability of the modeling approach to a full-scale industrial
design.
